---
title: "How wrong before it gets *wrong*? Radiocarbon sensitivity with OxCal and R."
subtitle: "Simulating oxcal models"
author: "Pete Jacobsson"
output: html_notebook
---

```{r setup, include=FALSE}
library(tidyverse)

```

## Introduction
This notebook contains three functions used to generate simulated OxCal model scripts, as well as usage examples in mass simulation contexts. For more details on the project background refer to *this medium blog*. For more details on OxCal, Bayesian chronological modelling and radiocarbon dating in general refer to *oxcal website* and the links therein.

The simulation functions are:
*  select_dates()
*  simulate_wmds()
*  simulate_seqs()

First of these functions is an auxiliary module function to run with simulate_wmds(). The latter two functions generate scripts that can be passed to OxCal to run randomized simulated Bayesian chronological models. simulate_wmds() simulates wiggle-match dates (wmds; REF WENINGER ET AL.) and simulate_seqs() simulates uniform bound sequences of one or more phases.

simulate_wmds and simulate_seqs() take on the following arguments:
* target (simulate_wmds) or tsb (target start boundary; simulate_seqs): provides a target date for the simulation. In case of wmds that will be the date of the outermost sample, while for sequences that will be the start boundary of a sequence. **Assumes that cal BP values are used**
* span: provides the total time over which a given timber grew, or deposition happened.
* n_samples: the number of samples passed to the simulated model. These will be placed equidistantly for wmds and in randomized locations for sequence models.
* phases (sequence models only): how many synchronous phases there are in a given sequence model.
* error: the magnitude of measurement errors associated with the simulated samples.
* offset: magnitude of systematic offset (in 14C yrs). This argument allows to simulate models with sampels that suffer from a systematic offset.
* offset_sigma: standard deviation of the systematic offset. This argument randomizes the systematic offset by modelling the offset as a normal distribution with mean = offset and standard deviation = offset_sigma. 

Note that the simulation functions provided do not take the IntCal predictive envelope into account - to simulate this with no systematic offset, set offset=0 and offset_sigma to the typical local value of the calibration curve for the period in which you're working.


These functions assume that you are working with an interpolated calibration curve. This can be generated using OxCal, or a relevant programming language. An OxCal generated interpolated IntCal20 for 0 - 12310 cal BP is provided.
```{r}
intcal_20 <- read_csv("IntCal20Interpolated.csv")
```

### Planned improvement
- Revise the offset_sigma argument, so that, if set to NA, the functions use IntCal predictive envelope for sample randomization (thus cloning OxCal's R_Simulate() function)
- Cross-test the scraping function and add it here.



## simulate_wmds() and select_dates()
```{r}
# select_dates--------------------------------------------------------------------- 
#A support function enabling date selection in automated generation of WMDs

select_dates <- function(target, span, n_samples) {
  #Takes in information on the target date, the span to be covered by the sim WMD and the number of Samples
  #Returns a vector of target dates for the wiggle-match
  #Assumes Span >= NSamples and all inputs should be integers

  #Original function used in project - produces symmetric distribution of simulation dates.
  #See below for a faster (and more obvious) function recommened for future application

  wmd_dates <- target

  for (i in 1:(n_samples-2)) {
    wmd_dates <- c(wmd_dates, wmd_dates[i] + (span/(n_samples-1)))
  }

  wmd_dates <- c(wmd_dates, target+span-1)
  wmd_dates <- floor(wmd_dates)
  wmd_dates

}

# select_dates <- function(target, span, n_samples) {
#   wmd_dates <- seq(from = target, to = target + span -1, length.out = n_samples)
#   wmd_dates <- floor(wmd_dates)
#   wmd_dates
# } #a much simpler alternative, though fails to pra symmetric selection


# simulate_wmds---------------------------------------------
#This function:
#simulates aseries of 14C measurements consistent with specification
#prints then to a format consistent with OxCal D_Sequence() function

simulate_wmds <- function(target, 
                           span, 
                           n_samples, 
                           error, 
                           offset = 0, offset_sigma =0, 
                           curve_to_oxcal = NA, 
                           cal_curve = intcal_20
                           ) {
  #The function takes in a series of parameters specifying:
  #wiggle-match target date, structure, offsets and calibration curve
  #The function returns a data frame that is consistent with OxCal input formatting
  #Function assumes that n_samples !> span, n_samples>=2 and that oxcal will recognize the calibration curve
  #span and n_samples need to be integers
  
  #Select target dates.
  if(n_samples == 2){sims_dates <- c(target, target+span-1)}else{
    sims_dates <- select_dates(target, span, n_samples)}##Use the Date Selector function to get the right cal BP values for the WMD
  
  #Extract the gaps between the measurements
  gaps <- rev(sims_dates[2:n_samples] - sims_dates[1:(n_samples-1)])
  
  sims_dates <- cal_curve[cal_curve$CalBP %in% sims_dates,]##Expand from cal BP Dates to All relevant data
  
  sims_ages <- rnorm(n_samples, sims_dates$BP, error) + offset ## Get the 14C Ages 
  sims_ages <- rnorm(n_samples, sims_ages, offset_sigma)
  sims_ages <- rev(sims_ages) # Consider piping in revision
  
  sims_ages
  
  if (is.na(curve_to_oxcal)) {
    open_wm <- paste(c('Plot(){D_Sequence( T', target, '){R_Date('), 
                     collapse="")} else {
      open_wm <- paste(c('Options(){Curve = "', curve_to_oxcal  ,'";};Plot(){D_Sequence( T', target, '){R_Date('), 
                       collapse="") 
    } #In a future revision consider replacing with str_c()
  commands <- c(open_wm, rep("R_Date(", n_samples-1))
  closing_statement <- c()
  for(i in 1:(n_samples-1)){
    closing_statement <- c(closing_statement,paste(c(",", error,");Gap(", gaps[i],");"), 
                                                collapse = ""))} #definite vectorization candidate
  close_wm <- paste(c(",", error,");};};"), collapse="")
  closing <- c(closing_statement, close_wm)
  
  oxcal_input <- data.frame(commands, sims_ages, closing)
  names(oxcal_input) = NULL
  rownames(oxcal_input) = NULL
  oxcal_input
  
  
  
    
}

#Sample usage
simulate_wmds(250, 30, 10, 20)
```

## simulate_seqs()
```{r}
#simulate_seqs-----------------------------------------------------------
#This function takes on a series of arguments to produce 
#a series of simulated 14C measurements printed into a model scripts consistent with oxcal's Sequence() model



simulate_seqs = function(tsb,                              #start date for the simulated sequence
                         span,                             #span of the simulated sequence
                         n_samples,                        #number of simulated measurements
                         phases,                           #number of phases within the simulated sequence
                         error,                            #simulated measurement error
                         offset = 0, offset_sigma = 0,     #simulated systematic offset and standard deviation of that offset
                         curve_to_oxcal = NA,              #specify curve for oxcal
                         cal_curve = intcal_20){            #specify from which curve to simulate
  
  teb <- tsb - span                                        #define sequence end boundary
  curve <- subset(cal_curve, CalBP >= teb &  CalBP <= tsb)
  
  sim_dates <- sample(curve$CalBP, n_samples, replace = TRUE)
  curve <- curve[curve$CalBP %in% sim_dates,]               #here we select the calibration curve values that will be used in the simulation
  sim_ages <- rep(NA, n_samples)
  for (i in 1:length(sim_dates)) {
    sim_ages[i] <- rnorm(1, 
                         curve[curve$CalBP %in% sim_dates[i],]$BP, 
                         error) + 
      offset
    sim_ages[i] <- rnorm(1, 
                         sim_ages[i], 
                         offset_sigma)
  }                                                         #run the 14C simulation. Consider vectorizing. Consider building offset randomization into a single line.
  sims_df <- data.frame(sim_dates, sim_ages)
  sims_df <- sims_df[rev(order(sims_df$sim_dates)),]        #assemble into data frame and re-arrange to older first (consistent with oxcal)
  
  phases_df <- data.frame(phase_number = seq(1, phases), 
                          sims_in_phase = rep(1, phases)) ## Set up DF tracking how many sims per each phase
  
  if (n_samples > phases) {
    for (i in 1:(n_samples - phases)) {
      to_add <- sample(phases, 1)
      phases_df$sims_in_phase[to_add] <- phases_df$sims_in_phase[to_add] + 1
    }
  }                                                         #distribute the number of measurements across multiple phases. Consider developing a vectorized approach.
  
  sims_cumulative <- c(phases_df$sims_in_phase[1])
  if (phases > 1) {
    for (i in 2:phases) {
      sims_cumulative <- c(sims_cumulative, 
                          sims_cumulative[i-1] + phases_df$sims_in_phase[i])
    }
  }
  
  phases_df <- cbind(phases_df, 
                    sim_phase_start = sims_cumulative - phases_df$sims_in_phase + 1,
                    sims_cumulative)                       #thus bringing together a DF saying how many samples per phase and cumulative numbers
  

  
  oxcal_input <- data.frame()
  for (i in 1:phases) {
    commands <- c('Phase(){R_Date(', 
                 rep('R_Date(', phases_df$sims_in_phase[i]-1))
    sim_ages <- sims_df$sim_ages[phases_df$sim_phase_start[i] : 
                                  phases_df$sims_cumulative[i]]
    closing_statement <- c(rep(paste(c(',', error, ');'), collapse = ""), phases_df$sims_in_phase[i]-1), 
                         paste(c(',', error, ');};'), collapse = ""))
    add_to_input <- data.frame(commands, sim_ages, closing_statement)
    oxcal_input <- rbind(oxcal_input, add_to_input)
  }                                                        #this sets up a df with all the phases in order, OxCal readable. 
                                                           #Uses a phase-by-phase approach.
                                                           #If possible, simplify.
  
  if (is.na(curve_to_oxcal)) {
    opening_commands <- paste(c('Plot(){Sequence(){Boundary("TSB', tsb, '");'), collapse = "")
  } else {
    opening_commands <- paste(c('Options(){Curve = "', curve_to_oxcal  ,
                               '";};Plot(){Sequence(){Boundary("TSB', tsb, '");'), collapse = "")
  }                                                       #this gets the opening commands for model script allowing for varying the calibration curve used
  
  closing_statement <- paste(c('Boundary("TEB', teb, '");};};'))   #this gets the closing commands for model script
  
  oxcal_input$commands <- as.character(oxcal_input$commands)
  oxcal_input$closing_statement <- as.character(oxcal_input$closing_statement)  #de-factorize
  
  oxcal_input[1,1] <-  paste(c(opening_commands, oxcal_input[1,1]), collapse = "")
  oxcal_input[n_samples, 3] <- paste(c(oxcal_input[n_samples, 3], closing_statement), collapse = "")  ##Add the Model opening and ending
  
  names(oxcal_input) = NULL
  rownames(oxcal_input) = NULL
  
  oxcal_input
}

#sample usage
simulate_seqs(2500, 30, 20, 2, 20)
```

## Usage in mass simulation.
The primary intent behind these functions is as components in mass simulation. For this purpose the workflow is as follows:
1. Create the table of values to be passed to the fucntions (a data frame). For parsing and checking down the line it is also useful to assign each simulation a unique number. 
2. Create names for the indivudal models (a vector).
3. Iterate through the table and the model names, passing the simulation outputs to OxCal.

Framework for passing the simulations to OxCal from R is copied of Richard Telford's blog (*LINK HERE*)

```{r, eval=FALSE}
#simulating 10000 wmds from 2300 - 2600 cal BP.

sample_wmd_params <- data.frame(sim_number = seq(1, 10000, 1), 
                                #unique simulation numbers
                                targets = sample(seq(2300, 2600, by=1), 10000, replace = T), 
                                #set target values to 2300 - 2600 cal BP
                                span = sample(seq(25, 100, by=1), 10000, replace = T),
                                #set span of the wmds to between 25 and 100 years
                                n_samples = sample(seq(2, 25, by=1), 10000, replace = T),
                                #set the number of samples in the wmds between 2 and 25
                                errors = sample(seq(8, 32, by=1), 10000, replace = T),
                                #set the measurement errors to between 8 and 32
                                offset = runif(10000, -50, 50),
                                #set offsets to between negative 50 and positive 50 14C years
                                offset_sigma = runif( 100000, 0.001, 26.5)
                                #set offset_sigma to between 0.01 and 26.5)

write.csv(sample_wmd_params, "sample_wmd_params.csv")
#save the generated parameters

sample_wmd_input_names <- rep("insert_file_path/sample_simulation", 10000) %>%
  str_c(seq(from = 1, to = 10000))
#running the simulations in a different directory prevents cluttering your r project directory


for (i in 1:10000) {
  write.table(simulate_wmds(target = sample_wmd_params$targets[i],
                            span = sample_wmd_params$span[i],
                            n_samples = sample_wmd_params$n_samples[i],
                            error = sample_wmd_params$errors[i],
                            offset = sample_wmd_params$offset[i],
                            offset_sigma = sample_wmd_params$offset_sigma[i]
                            ), #This generates the OxCal input script
              file  = str_c(sample_wmd_input_names[i], ".input"), 
              row.names = FALSE,
              quote = FALSE) #Inclusion of row names and quotes will result in the script becoming unreadable by OxCal.
  system(paste("your_oxcal_dir/bin/OxCalWin.exe", str_c(sample_wmd_input_names[i], ".input")))
  strip_oxcal_js(sample_wmd_input_names[i]) # Strip oxcal output .js file from the calibration curve data to save disk space.
}

```


